---
title: "Using the Crisp Template"
author: "Zhenguo Zhang"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using the Crisp Template}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The `crisp` template is a clean, modern resume template with customizable typography and layout options. It provides a clean, professional layout with tighter margins and spacing for creating concise resumes. This vignette demonstrates how to use the template with various entry types.

## Template Features

- Monochrome output (no colors, except for the section titles)
- Customize the color of first 3 characters in section title with 'headcolor'
- Compact layout with tighter margins
- Flexible entry types (detailed, brief, honors, skills)
- Customizable spacing after section title using 'spacing_after_section_title'
- Header with inline contact information
- Choose from the font family: Roboto, Times New Roman
- All other header options in awesomecv, such as `address`, `email`, etc.

## Basic Setup

First, load the required packages:

```{r setup}
library(vitae)
library(tibble)
library(dplyr)
```

## Creating a Resume

### Document Header

Create your resume by setting up an R Markdown document with the following YAML header:

```yaml
---
title: "Your Name"
author: Your Name
date: "`r Sys.Date()`"
output: vitae::crisp
position: "Data Scientist"
address: "123 Main Street, City, Country"
phone: "+1 234 567 8900"
email: "email@example.com"
github: "fortune9"
linkedin: "username"
headcolor: "0000FF"      # Blue color for the 3 characters in section title
spacing_after_section_title: "-1mm"  # Compress section spacing
---
```

### Using Different Entry Types

#### Detailed Entries

Use `detailed_entries()` for experience sections like work history or education:

```{r eval=FALSE}
tribble(
  ~title, ~unit, ~loc, ~dates, ~details,
  "Ph.D.", "University of Statistics", "California", "2020-2024",
  c("Thesis: Statistical Analysis of Data",
    "Published 3 papers in top journals",
    "Received Excellence Award"),
  "M.Sc.", "College of Data", "New York", "2018-2020",
  c("Focus on Machine Learning",
    "GPA: 4.0/4.0")
) %>%
  detailed_entries(
    what = title,
    when = dates,
    with = unit,
    where = loc,
    why = details
  )
```

#### Brief Entries

Use `brief_entries()` for shorter items like publications or certifications:

```{r eval=FALSE}
tribble(
  ~role, ~year, ~inst,
  "Best Paper Award", "2023", "Statistics Conference",
  "R Programming Certificate", "2022", "DataCamp"
) %>%
  brief_entries(
    what = role,
    when = year,
    with = inst
  )
```

#### Honor Entries

The new `honor_entries()` function is perfect for achievements and awards:

```{r eval=FALSE}
tribble(
  ~award, ~year, ~org, ~location,
  "Outstanding Researcher", "2023", "Science Foundation", "National",
  "Teaching Excellence", "2022", "University", "Department"
) %>%
  honor_entries(
    what = award,
    when = year,
    with = org,
    where = location
  )
```

#### Skill Entries

Use `skill_entries()` to showcase your expertise:

```{r eval=FALSE}
tribble(
  ~category, ~details,
  "Programming", "R, Python, SQL, Julia",
  "Statistics", "Regression, Time Series, Machine Learning",
  "Languages", "English (Native), Spanish (Fluent)"
) %>%
  skill_entries(
    type = category,
    details = details
  )
```

## Complete Example

Here's a minimal working example combining all entry types:

```{r eval=FALSE}
---
title: "Professional Resume"
author: Jane Doe
output: vitae::crisp
position: "Data Scientist"
email: "jane.doe@email.com"
github: "janedoe"
spacing_after_section_title: "-1mm"
headcolor: "00FF00" # green
---

# Education

```{r}
tribble(
  ~degree, ~uni, ~loc, ~dates, ~details,
  "Ph.D.", "Stats University", "USA", "2020-2024", c("Thesis on Advanced Analytics"),
  "B.S in Biology", "Biology University", "USA", "2016-2020", c("Thesis on Biology")
) %>% 
  detailed_entries(
    what = degree,
    when = dates,
    with = uni,
    where = loc,
    why = details
  )
```

# Skills

```{r}
tribble(
  ~category, ~details,
  "Analysis", "Statistics, Machine Learning, Data Mining",
  "Programming", "R, Python, SQL",
  "Reserch", "Biology, Genetics",
  "Visualization", "Rmarkdown, streamlit, shiny",
  "DevOps", "Github, AWS, GCP, docker, nextflow"
) %>%
  skill_entries(
    type = category,
    details = details
  )
```

# Honors

```{r}
tribble(
  ~award, ~year, ~org, ~loc,
  "Best Paper1", "2023", "Conference4", "International",
  "Best Paper2", "2022", "Conference3", "International",
  "Best Paper3", "2021", "Conference2", "International",
  "Best Paper4", "2020", "Conference1", "International"
) %>%
  honor_entries(
    what = award,
    when = year,
    with = org,
    where = loc
  )
```

# Publications

```{r}
tribble(
  ~title, ~year, ~journal,
  "Amazing Research Paper", "2023", "Journal of Data Science",
  "Bioinformatics Paper", "2022", "Journal of Bioinformatics",
  "Genetics Paper","2020","Journal of Genetics"
) %>%
  brief_entries(
    what = title,
    when = year,
    with = journal
  )
```
```

## Customization Tips

1. **Section Spacing**: Adjust vertical space between sections using `spacing_after_section_title` in YAML header
2. **Monochrome Output**: Set `headcolor: "000000"` in YAML header
3. **Entry Spacing**: Each entry type has built-in spacing optimized for compact layout
4. **Protection**: Use `.protect = TRUE` (default) to handle LaTeX special characters

## Common Issues and Solutions

1. **LaTeX Errors**: Ensure you have a complete TeX distribution installed
2. **Font Issues**: The template uses specific fonts that should be installed with the package
3. **Spacing**: Fine-tune using `spacing_after_section_title` and LaTeX spacing commands if needed

## Conclusion

The `crisp` template provides a professional, compact layout suitable for most resume needs. By combining different entry types (`detailed_entries`, `brief_entries`, `honor_entries`, and `skill_entries`), you can create a well-structured and comprehensive resume.